---
title: "Lab 02 - Plastic waste"
author: "Ella Beiser"
date: "January 27, 2022"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read_csv("data/plastic-waste.csv")
```

## Exercises

```{r}
plastic_waste <- read_csv("data/plastic-waste.csv")
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)
plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)
```
Trinidad and Tobago produce the most waste per capita per day in the world according to a report by Forbes. Additionally, their waste is improperly managed and likely to end up in the ocean.

### Exercise 1

The density plot below shows that the majority of countries in Africa have very low plastic waste per capita with the majority producing 0.1 kilograms per capita. In contrast, Europe has a wide range of plastic waste production with the majority ranging from 0.15-0.27 kilograms per capita. The plot also indicates that each continent has a number of outliers who produce significantly more plastic waste per capita than the rest.  

```{r plastic-waste-continent}
#Below is a histogram of plastic waste faceted
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram()
#Below is a density plot with color!
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density(alpha = 0.7)
```

### Exercise 2
```{r plastic-waste-density}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density(alpha = 0.4) +
  xlim(0, 1)
```

### Exercise 3

The color and fill of the curves are defined by the mapping aesthetic because they are dependent on the data and it is not consistent across the board. In contrast, the alpha level is a characteristic of the plotting geom because the alpha is a consistent level for the entire plot. 

```{r}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_boxplot()
```

### Exercise 4

The box plot automatically calculates summary statistics like the mean, outliers etc and indicates them on the graph. Whereas in the violin plot, the plot shows approximately where the mean would be but it is less clear. 
```{r plastic-waste-violin}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()
```

### Exercise 5

This scatterplot shows that as plastic waste increases, mismanaged plastic waste also increases. The rate at which it increases varies significantly. There is also an outlier with significant plastic waste but lower mismanaged plastic waste per capita. 
```{r plastic-waste-mismanaged}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     y = mismanaged_plastic_waste_per_cap)) +
  geom_point()
```

### Exercise 6

With color, it becomes more evident the relationship between plastic waste and mismanaged plastic waste per capita. For Africa it appears as if there is a strong positive correlation between the two. The rest of the continents are less clear. 
```{r plastic-waste-mismanaged-continent}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     y = mismanaged_plastic_waste_per_cap, color = continent, fill = continent)) +
  geom_point()
```

### Exercise 7

The relationship between the coastal populations and plastic waste per capita appear to be more linearly associated. 
```{r plastic-waste-population-total}
ggplot(data = plastic_waste, 
       mapping = aes(x = total_pop, 
                     y = plastic_waste_per_cap)) +
  geom_point()
```

```{r plastic-waste-population-coastal}
ggplot(data = plastic_waste, 
       mapping = aes(x = coastal_pop, 
                     y = plastic_waste_per_cap)) +
  geom_point()
```

### Exercise 8
From this plot we can interpret that until the coastal population proportion reaches 0.75 the plastic waste per capita increases with it but after that point it levels off and slowly declines. Additionally, the vast majority of countries have a coast population proportion below 1.0. 

```{r recreate-viz}
ggplot(data = plastic_waste) +
  geom_point(mapping = aes(x = coastal_pop/plastic_waste_per_cap, 
                     y = plastic_waste_per_cap, color = continent)) +
  geom_smooth(mapping = aes(x = coastal_pop/plastic_waste_per_cap, 
                     y = plastic_waste_per_cap))+ ylim(0, 0.7) + xlim(0, 2.0e+07) + labs(x = "Coastal population proportion (Coastal / total population)", y="Plastic waste per capita", title = "Plastic waste vs Coastal population proportion", subtitle = "by continent", color = "Continent") 

```
